Racial and Ethnic Differences in ETI: Outcomes

### 2024-12-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
adjust for prior modulator, pseudomonas, etc.

```{r}
# Create change in slope variable
encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_change_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_change_years) <- "Change in Slope"
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
  Insurance = "Private"
))
pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

>>>>

```{r}
# Create change in slope variable
encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_change_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_change_years) <- "Change in Slope"
# Prior modulator use
prior_mods <- encounter %>%
  select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
  filter(Days < 0) %>%
  group_by(eDWID) %>%
  summarise(prior_modulator = any(vx661comb == 1) | any(vx445comb == 1) |
    any(Vx770 == 1) | any(Vx809comb))
prior_mods$prior_modulator <-
  fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
prior_mods$prior_modulator <- factor(prior_mods$prior_modulator,
  levels = c(T, "No"),
  labels = c("Yes", "No")
)
encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
# On enzymes for pancreatic insufficiency
encounter$isOnEnzymes <-
  factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
# Pseudomonas
encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
encounter$pseudomonasaeruginosa <- factor(encounter$pseudomonasaeruginosa,
  levels = 0:1, labels = c("No", "Yes")
)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
  Insurance = "Private", prior_modulator = "Yes",
  isOnEnzymes = "Yes", pseudomonasaeruginosa = "No"
))
pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt(caption = "Predicted Values at Various Timepoints") %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
    age_group + Insurance + prior_modulator + isOnEnzymes +
    pseudomonasaeruginosa + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race + Insurance +
    prior_modulator + isOnEnzymes + pseudomonasaeruginosa + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race + slope_change_years * Race +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_other_pre <- glht(
  change_mod,
  "Years + `Years:RaceOther` = 1.19547"
)
white_vs_other_pre <- tidy(summary(white_vs_other_pre))
white_vs_other_pre$contrast <- "Other vs. White Pre-ETI Slope"
white_vs_other_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_other_post <- glht(
  change_mod,
  "Years + `Years:RaceOther` + slope_change_years + `RaceOther:slope_change_years` = 2.738885"
)
white_vs_other_post <- tidy(summary(white_vs_other_post))
white_vs_other_post$contrast <- "Other vs. White Post-ETI Slope"
white_vs_other_post %>%
  gt() %>%
  fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race + age_group +
    Insurance + prior_modulator + isOnEnzymes + pseudomonasaeruginosa +
    (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 3)
# Linear hypotheses
white_vs_black_pre <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` = 1.195465"
)
white_vs_black_pre <- tidy(summary(white_vs_black_pre))
white_vs_black_pre$contrast <- "Black vs. White Pre-ETI Slope"
white_vs_black_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_black_post <- glht(
  change_mod,
  "Years + `Years:RaceBlack or African American` + slope_change_years + `RaceBlack or African American:slope_change_years` = 2.738885"
)
white_vs_black_post <- tidy(summary(white_vs_black_post))
white_vs_black_post$contrast <- "Black vs. White Post-ETI Slope"
white_vs_black_post %>%
  gt() %>%
  fmt_number(decimals = 3)

white_vs_hispanic_pre <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` = 1.19547"
)
white_vs_hispanic_pre <- tidy(summary(white_vs_hispanic_pre))
white_vs_hispanic_pre$contrast <- "Hispanic vs. White Pre-ETI Slope"
white_vs_hispanic_pre %>%
  gt() %>%
  fmt_number(decimals = 3)
white_vs_hispanic_post <- glht(
  change_mod,
  "Years + `Years:RaceHispanic or Latino` + slope_change_years + `RaceHispanic or Latino:slope_change_years` = 2.738885"
)
white_vs_hispanic_post <- tidy(summary(white_vs_hispanic_post))
white_vs_hispanic_post$contrast <- "Hispanic vs. White Post-ETI Slope"
white_vs_hispanic_post %>%
  gt() %>%
  fmt_number(decimals = 3)
